---
title: "RPQ analysis - Mindset"
author: "Yizhou Yu"
date: "NA"
output: 
 html_document: 
    toc: true
    toc_float: true
---

The Rivermead Post-concussion Symptoms Questionnaire (RPQ) is a widely used tool for assessment of the presence and severity of various post-concussion symptoms. The questionnaire was first published in 1995 to assess patients' psychosocial functioning post-concussion. It contains 16 questions targeting physical, cognitive and behavioural domains. It was modified by Eyres et al. is 2005 into RPQ Modified Scoring System (RPQ 13/ RPQ-3)

Workflow: 
- Descriptive statistics
- Do individuals with recent concussions score high on RPQ3 and whether anyone with less recent concussion symptoms?
- Identify the main questions involved in the RPQ  (13 and 3)
- See whether fewer questions in addition to the Mscores could predict the RPQ parameters. 

## Data curation
load libraries
```{r, message=FALSE}
library(tidyverse)
library(ggplot2)
library(pdp)
library(rpart.plot)
library(caret)
library(dplyr)
```


## Data curation
load data<br>
Note: in the metadt, when use == 2, the variables will have multiple components
```{r}
metadt = read.csv("metadt.csv")
metadt = subset(metadt, use == 1)
dt_raw = read.csv("dt/NK_14_04_2021.csv")
dt <- dt_raw[, metadt$varname]
```

Add additional data: 

Add RPQ questions: RPQ_Q1-16

"RPQ_Q1_CORRECTED"                        "RPQ_Q2_CORRECTED"
"RPQ_Q3_CORRECTED"                        "RPQ_Q4_CORRECTED"
"RPQ_Q5_CORRECTED"                        "RPQ_Q6_CORRECTED"
"RPQ_Q7_CORRECTED"                        "RPQ_Q8_CORRECTED"
"RPQ_Q9_CORRECTED"                        "RPQ_Q10_CORRECTED"
"RPQ_Q11_CORRECTED"                       "RPQ_Q12_CORRECTED"
"RPQ_Q13_CORRECTED"                       "RPQ_Q14_CORRECTED"
"RPQ_Q15_CORRECTED"                       "RPQ_Q16_CORRECTED"                      
"RPQ_TOT_CORRECTED"                       "RPQ13_CORRECTED"  

```{r}
rpq_q_list = c()

for (i in 1:16){
  rpq_question = paste("RPQ_Q", i, "_CORRECTED", sep = "")
  rpq_q_list = c(rpq_q_list, rpq_question)
}
rpq_dt = subset(read.csv("dt/NK_14_04_2021.csv"), select = rpq_q_list)
head(rpq_dt,2)
```

```{r}
subset(read.csv("metadt.csv"), use == 2)
```

```{r}
dt_mci = dt_raw %>% select(starts_with("MCI_"))
dt_neuro = dt_raw %>% select(starts_with("neuro_"))
dt_RPQ = dt_raw %>% select(starts_with("RPQ"))
dt_neuro$neuro_TOT = rowSums(dt_neuro)
dt_MH = dt_raw %>% select(starts_with("MH_"))
dt_MH$MH_TOT = rowSums(dt_MH)
```

```{r}
#dt = cbind(dt,dt_mci) # use MCI_TOT instead
dt$neuro_TOT = dt_neuro[,"neuro_TOT"]
dt$MH_TOT = dt_MH[,"MH_TOT"]
```

exclude scores
```{r}
exclude_vars <- names(dt) %in% c("CAGE_TOT", #alcohol
                                    "GAD7_TOT", #anxiety
                                    "PHQ2_TOT", #depression
                                    "PHQ9_TOT") #depression
dt_noS = dt[,!exclude_vars]
head(dt_noS)
```
## Descriptive data

```{r}
num_graph = dt_noS %>%
  keep(is.numeric)
discrete_graph = dt_noS[ , !(names(dt_noS) %in% colnames(num_graph))]

discrete_graph_gather = discrete_graph %>%
  gather() 

ggplot(discrete_graph_gather, aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram(stat = "count",fill="#939598",color="#939598") +
    theme(
      text = element_text(colour = "#231F20"),
      strip.background =element_rect(fill="transparent"),
      axis.text.x = element_text(angle = 90,color = "#231F20"),
      #axis.text.x = element_text(colour = "#231F20"),
      axis.text.y = element_text(colour = "#231F20"),
      panel.background = element_rect(fill = "transparent"), # bg of the panel
      plot.background = element_rect(fill = "transparent", color = NA), # bg of the plot
      panel.grid.major = element_blank(), # get rid of major grid
      panel.grid.minor = element_blank())# get rid of minor grid coord_flip()
ggsave('fig/dt_distribution_discrete.pdf', width = 8, height = 10)
```

```{r}
num_graph %>%
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram(fill="#939598",color="#939598") + 
    theme(
      text = element_text(colour = "#231F20"),
      axis.text.x = element_text(colour = "#231F20"),
      axis.text.y = element_text(colour = "#231F20"),
      strip.background =element_rect(fill="transparent"),
      panel.background = element_rect(fill = "transparent"), # bg of the panel
      plot.background = element_rect(fill = "transparent", color = NA), # bg of the plot
      panel.grid.major = element_blank(), # get rid of major grid
      panel.grid.minor = element_blank())# get rid of minor grid
ggsave('fig/dt_distribution_continuous.pdf', width = 6.3, height = 7)
```

```{r}
nrow(na.omit(dt_noS))
```

## Recent concussion and RPQ questions

Reformat the data: I assume that Less than 1 year => recent, and more than 1 year => previous
```{r}
concussion_subanalysis = subset(dt_noS, select = c(age,mostRecentConcussion,
                                                   RPQ3_CORRECTED,
                                                   RPQ13_CORRECTED))
concussion_subanalysis_dt = na.omit(concussion_subanalysis)
concussion_subanalysis_dt$recent_concussion = with(concussion_subanalysis_dt, ifelse(mostRecentConcussion %in% 
                                                 c("New_User_ConcussionC_1", 
                                                   "New_User_ConcussionC_2"), TRUE, FALSE))
```

Scale: RPQ3 is out of 12, and RPQ13 is out of 52
```{r}
concussion_subanalysis_dt$norm_RPQ3 = concussion_subanalysis_dt$RPQ3_CORRECTED/12
concussion_subanalysis_dt$norm_RPQ13 = concussion_subanalysis_dt$RPQ13_CORRECTED/52
```

```{r}
summary(norm_RPQ3_lm <- lm(data = concussion_subanalysis_dt, norm_RPQ3 ~ recent_concussion + age))
wilcox.test(concussion_subanalysis_dt$norm_RPQ3 ~ concussion_subanalysis_dt$recent_concussion)
confint(norm_RPQ3_lm)
```
```{r}
wilcox.test(concussion_subanalysis_dt$norm_RPQ13 ~ concussion_subanalysis_dt$recent_concussion)
```
```{r}
prev_conc_lm = lm(data = concussion_subanalysis_dt, norm_RPQ13 ~ age + recent_concussion)
summary(prev_conc_lm)
```

```{r}
rec_conc_lm = lm(data = concussion_subanalysis_dt, norm_RPQ3 ~ age + recent_concussion)
summary(rec_conc_lm)

ggplot(data = concussion_subanalysis_dt, aes(x=recent_concussion, y=norm_RPQ3)) + 
  geom_boxplot()+ theme_classic()+
  theme(
    panel.background = element_rect(fill = "transparent"), # bg of the panel
    plot.background = element_rect(fill = "transparent", color = NA), # bg of the plot
    panel.grid.major = element_blank(), # get rid of major grid
    panel.grid.minor = element_blank(), # get rid of minor grid
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent") # get rid of legend panel bg
  )
ggsave("fig/RPQ3_recent_concussion_validation.pdf", width = 2, height = 2,  bg = "transparent")
```

Conclusion: a recent concussion is associated with higher RPQ3 score (estimate = 0.1057319),
but a previous concussion is not associated with a higher RPQ13 score, indicating that perhaps some people might be recovering well, while others are not. It might be interesting to explore this further.

### Further analyses on RPQ3 

I did not pursue further tests only on RPQ3 because there were only 22 individuals

```{r}
rpq3_analysis_subset = dt

rpq3_analysis_subset$recent_concussion = with(rpq3_analysis_subset, ifelse(mostRecentConcussion %in% 
                                                 c("New_User_ConcussionC_1", 
                                                   "New_User_ConcussionC_2"), TRUE, FALSE))
rpq3_analysis_subset = subset(rpq3_analysis_subset, recent_concussion == TRUE)
nrow(rpq3_analysis_subset)

# run a for loop to screen for all confounders one by one
variable_list = colnames(rpq3_analysis_subset)
variable_list = variable_list[variable_list != "age"]
variable_list = variable_list[variable_list != "hadConcussion"]
variable_list = variable_list[variable_list != "RPQ3_CORRECTED"]

Estimate=c()
Std..Error=c()
t.value=c()
Pr...t..=c()
var_names=c()

sumamry_list = data.frame(Estimate,Std..Error,t.value,Pr...t..,var_names)
for (var in 1:length(variable_list)){
  var_name = variable_list[var]
  lm_formula = as.formula(
    paste("RPQ3_CORRECTED",paste(c("age" , var_name), collapse = " + "),sep = " ~ ")
  )
  # print(lm_formula)
  summary_var = summary(rpq3_subanalysis_lm <- lm(formula = lm_formula, data = rpq3_analysis_subset))
  summary_df = data.frame(summary_var$coefficients)
  summary_df$var_names = row.names(summary_df)
  summary_df_clean = subset(summary_df, !(var_names %in% 
                                            c('(Intercept)','age')))
  sumamry_list = rbind(sumamry_list,summary_df_clean)
}

```

```{r}
sumamry_list_significant = subset(sumamry_list, Pr...t.. <= 0.05)
sumamry_list_significant
```

Having a neurological condition seem to increase the RPQ score

```{r}
# summary(rpq3_subanalysis_lm_neuro <- lm(formula = (RPQ3_CORRECTED)/12 ~ 
#                                     age + neuro_TOT, data = rpq3_analysis_subset))
rpq3_analysis_subset_plot = rpq3_analysis_subset
rpq3_analysis_subset_plot$neuro_condition = with(rpq3_analysis_subset_plot, ifelse(neuro_TOT >= 1, TRUE, FALSE))

summary(rpq3_subanalysis_lm_neuro <- lm(formula = (RPQ3_CORRECTED)/12 ~ 
                                    age + neuro_condition, data = rpq3_analysis_subset_plot))

ggplot(data = rpq3_analysis_subset_plot, aes(x=neuro_condition, y=RPQ3_CORRECTED/12)) + 
  geom_boxplot() + theme_classic()+
  theme(
    panel.background = element_rect(fill = "transparent"), # bg of the panel
    plot.background = element_rect(fill = "transparent", color = NA), # bg of the plot
    panel.grid.major = element_blank(), # get rid of major grid
    panel.grid.minor = element_blank(), # get rid of minor grid
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent") # get rid of legend panel bg
  )
ggsave("fig/RPQ3_subanalysis_neuro_condition.pdf", width = 2, height = 2,  bg = "transparent")
```

```{r}
confint(rpq3_subanalysis_lm_neuro)
length(rpq3_analysis_subset_plot$neuro_condition[rpq3_analysis_subset_plot$neuro_condition== TRUE])
```

8 have neurological condition and 22-8 = 14 do not. 

## Simplification of RPQ3 

```{r}
rpq3_subanalysis = cbind(subset(dt_noS, select = c(RPQ3_CORRECTED)),rpq_dt[,1:3])
```


Do a PCA for RPQ3
```{r}
RPQ3_pca = FactoMineR::PCA(rpq3_subanalysis[,-1], scale.unit = TRUE, ncp = 5, graph = TRUE)
```

```{r}
factoextra::fviz_pca_ind(RPQ3_pca,
             geom.ind = "point", # show points only (nbut not "text")
             ) + theme_classic()+
  theme(
    panel.background = element_rect(fill = "transparent"), # bg of the panel
    plot.background = element_rect(fill = "transparent", color = NA), # bg of the plot
    panel.grid.major = element_blank(), # get rid of major grid
    panel.grid.minor = element_blank(), # get rid of minor grid
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent") # get rid of legend panel bg
  )
ggsave("fig/RPQ3_PCA_general_graph.pdf", width = 3, height = 2,  bg = "transparent")
```

```{r}
factoextra::fviz_eig(RPQ3_pca, addlabels = TRUE, ylim = c(0, 50)) + theme_classic()
ggsave("fig/RPQ3_PCA_components.pdf", width = 3, height = 2,  bg = "transparent")
```

```{r}
factoextra::fviz_contrib(RPQ3_pca, choice = "var", axes = 1, top = 100) + theme_classic()
ggsave("fig/RPQ3_PCA_contribution.pdf", width = 3, height = 2,  bg = "transparent")
```

This shows that RPQ2 is the highest predictor of RPQ3. <br>Q2 is : Compared with before the accident, do you now (i.e., over the last 24 hours) suffer from Feelings of Dizziness

Q1 is Headaches and Q3 is Nausea


### RPQ - Q2 and RPQ3
```{r}
rpq3_validation = cbind(dt_noS,rpq_dt)
rpq3_validation.lm = lm(data = rpq3_validation,RPQ3_CORRECTED~RPQ_Q2_CORRECTED + age)
summary(rpq3_validation.lm)

```

```{r}
# rpq3_validation.lm

rpq3_validation.lm_dt = data.frame(cbind((cbind(Beta = coef(rpq3_validation.lm), 
                                                confint(rpq3_validation.lm))), p_value = summary(rpq3_validation.lm)$coefficients[,4]))

rpq3_validation.lm_dt$significance = "p-value > 0.05"
rpq3_validation.lm_dt$significance[rpq3_validation.lm_dt$p_value < 0.05] <- "p-value < 0.05"
rpq3_validation.lm_dt$name= row.names(rpq3_validation.lm_dt)
rpq3_validation.lm_dt = subset(rpq3_validation.lm_dt, name != "(Intercept)")
ggplot2::ggplot(rpq3_validation.lm_dt, 
       aes(x=name, y=Beta, color = p_value, fill = p_value)) + 
    geom_point(shape=21, size = 2.5) +
    geom_errorbar(aes(ymin=X2.5.., ymax=X97.5..),
                  width=0,                    # Width of the error bars
                  size = 1.2,
                  position=position_dodge(.9)) +
  scale_color_gradient(
  low = "#ED2024",
  high = "#939598") + 
  scale_fill_gradient(
  low = "#ED2024",
  high = "#939598") + 
  theme_classic() + 
  # geom_hline(yintercept = 1, linetype="dotted") +
  coord_flip()+ylab("Beta") + 
  xlab("") 
ggsave("fig/rpq3_question2_lm.pdf", width = 4, height = 1.5,  bg = "transparent")
```

## Simplification of RPQ13 

```{r}
prev_conc_lm = lm(data = concussion_subanalysis_dt, norm_RPQ13 ~ age + recent_concussion)
summary(prev_conc_lm)

ggplot(data = concussion_subanalysis_dt, aes(x=recent_concussion, y=norm_RPQ13)) + 
  geom_boxplot()+ theme_classic()+
  theme(
    panel.background = element_rect(fill = "transparent"), # bg of the panel
    plot.background = element_rect(fill = "transparent", color = NA), # bg of the plot
    panel.grid.major = element_blank(), # get rid of major grid
    panel.grid.minor = element_blank(), # get rid of minor grid
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent") # get rid of legend panel bg
  )
ggsave("fig/RPQ13_previous_concussion_validation.pdf", width = 2, height = 2,  bg = "transparent")
```

```{r}
rpq13_subanalysis = cbind(subset(dt_noS, select = c(RPQ13_CORRECTED)),rpq_dt[,4:16])

RPQ13_pca = FactoMineR::PCA(rpq13_subanalysis[,-1], scale.unit = TRUE, ncp = 5, graph = TRUE)
```

```{r}
factoextra::fviz_pca_ind(RPQ13_pca,
             geom.ind = "point", # show points only (nbut not "text")
             ) + theme_classic()+
  theme(
    panel.background = element_rect(fill = "transparent"), # bg of the panel
    plot.background = element_rect(fill = "transparent", color = NA), # bg of the plot
    panel.grid.major = element_blank(), # get rid of major grid
    panel.grid.minor = element_blank(), # get rid of minor grid
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent") # get rid of legend panel bg
  )
ggsave("fig/RPQ13_PCA_general_graph.pdf", width = 3, height = 2,  bg = "transparent")
```

```{r}
factoextra::fviz_eig(RPQ13_pca, addlabels = TRUE, ylim = c(0, 50))+ theme_classic()
ggsave("fig/RPQ13_PCA.pdf", width = 3, height = 2,  bg = "transparent")
```

```{r}
factoextra::fviz_contrib(RPQ13_pca, choice = "var", axes = 1, top = 100) + theme_classic() + 
  theme(axis.text.x = element_text(angle = 90))
ggsave("fig/RPQ3_PCA_contributions.pdf", width = 3, height = 4,  bg = "transparent")
```

This shows that RPQ2 is the highest predictor of RPQ3. <br>Q2 is : Compared with before the accident, do you now (i.e., over the last 24 hours) suffer from Feelings of Dizziness

Q1 is Headaches and Q3 is Nausea


### RPQ - Q9 and RPQ13
```{r}
rpq13_validation = cbind(dt_noS,rpq_dt)
rpq13_validation.lm = lm(data = rpq3_validation,RPQ13_CORRECTED~RPQ_Q9_CORRECTED + 
                           RPQ_Q12_CORRECTED + 
                           RPQ_Q11_CORRECTED + 
                           RPQ_Q6_CORRECTED +
                           RPQ_Q10_CORRECTED +
                           RPQ_Q7_CORRECTED +
                           RPQ_Q16_CORRECTED +
                           RPQ_Q8_CORRECTED)
summary(rpq13_validation.lm)

```

```{r}
rpq13_validation2 = cbind(dt_noS,rpq_dt)
rpq13_validation2.lm = lm(data = rpq13_validation,RPQ13_CORRECTED~
                           RPQ_Q9_CORRECTED)
summary(rpq13_validation2.lm)

```
I tested whether adding any other variable will improve the scoring: RPQ Q6 improves the R2 from 0.65 to 0.82, which is the largest increase.

RPQ Q6: Fatigue, tiring more easily
RPQ Q9: Frustrated
```{r}
rpq13_validation.lm = lm(data = rpq3_validation,RPQ13_CORRECTED~RPQ_Q9_CORRECTED + 
                           #RPQ_Q12_CORRECTED + 
                           #RPQ_Q11_CORRECTED + 
                           RPQ_Q6_CORRECTED #+
                           #RPQ_Q10_CORRECTED +
                           #RPQ_Q7_CORRECTED +
                           #RPQ_Q16_CORRECTED +
                           #RPQ_Q8_CORRECTED
                         +age
                         )
summary(rpq13_validation.lm)

```

```{r}
rpq13_validation.lm_df = data.frame(name=row.names(summary(rpq13_validation.lm)$coefficients),cbind((cbind(Beta = coef(rpq13_validation.lm), confint(rpq13_validation.lm))), p_value = summary(rpq13_validation.lm)$coefficients[,4]))

write.csv(rpq13_validation.lm_df,"dt_out/rpq13_validation_lm.csv")

rpq13_validation.lm_df = subset(rpq13_validation.lm_df, name != "(Intercept)")

ggplot(rpq13_validation.lm_df, 
       aes(x=reorder(name,Beta), y=Beta, color = p_value, fill = p_value)) + 
    geom_point(shape=21, size = 2.5) +
    geom_errorbar(aes(ymin=X2.5.., ymax=X97.5..),
                  width=0,                    # Width of the error bars
                  size = 1.2,
                  position=position_dodge(.9)) +
  scale_color_gradient(
  low = "#ED2024",
  high = "#939598") + 
  scale_fill_gradient(
  low = "#ED2024",
  high = "#939598") + 
  theme_classic() + 
  # geom_hline(yintercept = 1, linetype="dotted") +
  coord_flip()+ylab("Beta") + 
  xlab("") + 
  annotate("text", x = 2.5, y=1, label= "Adjusted R-squared:   0.82")

ggsave('fig/rpq13_validation_lm.pdf', width = 4, height = 1.5,  bg = "transparent")
```


```{r}
cor.test(rpq3_validation$RPQ_Q9_CORRECTED, rpq3_validation$RPQ_Q6_CORRECTED)
rpq13_plot = na.omit(subset(rpq3_validation, select = c(RPQ_Q6_CORRECTED,RPQ_Q9_CORRECTED)))
ggplot(data = rpq13_plot, aes(x = RPQ_Q6_CORRECTED, y = RPQ_Q9_CORRECTED)) + 
  geom_point() +
  geom_smooth(method = "lm") + 
  theme_classic()
```


## Modifiable factors

### Dementia
```{r}
dt_noS_dementia = subset(dt_noS, select = c(mostRecentConcussion,
                                           RPQ13_CORRECTED,
                                           # BM7 + 
                                           age,
                                           # ageStroopQscore + 
                                           # concussions,
                                           # covidPositive + 
                                           dementia
                                           # hadConcussion,
                                           # joke + 
                                           # memory + 
                                           # alcohol,
                                           # alcoholAmount,
                                           # smoking,
                                           # MCI_TOT
                                           ))
dt_noS_dementia = na.omit(dt_noS_dementia)
dt_noS_dementia$recent_concussion = with(dt_noS_dementia, ifelse(mostRecentConcussion %in% 
                                                 c("New_User_ConcussionC_1", 
                                                   "New_User_ConcussionC_2"), TRUE, FALSE))
dt_noS_dementia$dem_recoded = with(dt_noS_dementia, ifelse(dementia %in% 
                                                 c("ResponseNo", 
                                                   "ResponseDontKnow"), FALSE, TRUE))

dt_long_concuss = subset(dt_noS_dementia, recent_concussion == FALSE)

summary(rpq13_lm <-lm(data = dt_long_concuss, RPQ13_CORRECTED ~ 
                                           # BM7 + 
                                           age + 
                                           # ageStroopQscore + 
                                           # concussions,
                                           # covidPositive + 
                                           dementia# + 
                                           # hadConcussion,
                                           # joke + 
                                           # memory + 
                                           # alcohol * 
                                           # alcoholAmount + 
                                           # smoking + 
                                           # MCI_TOT 
                                           # neuro_TOT + 
                                           # MH_TOT
                                   ))

rpq13_lm_dt = data.frame(cbind((cbind(Beta = coef(rpq13_lm), confint(rpq13_lm))), p_value = summary(rpq13_lm)$coefficients[,4]))

write.csv(rpq13_lm_dt,"dt_out/rpq13_lm_dt_dementia.csv", row.names=TRUE)

```


### Smoking

```{r}
dt_noS_subset = subset(dt_noS, select = c(mostRecentConcussion,
                                           RPQ13_CORRECTED,
                                           # BM7 + 
                                           age,
                                           # ageStroopQscore + 
                                           # concussions,
                                           # covidPositive + 
                                           # dementia
                                           # hadConcussion,
                                           # joke + 
                                           # memory + 
                                           # alcohol,
                                           # alcoholAmount,
                                           smoking
                                           # MCI_TOT
                                           ))

dt_noS_subset$recent_concussion = with(dt_noS_subset, ifelse(mostRecentConcussion %in% 
                                                 c("New_User_ConcussionC_1", 
                                                   "New_User_ConcussionC_2"), TRUE, FALSE))

dt_long_concuss = subset(dt_noS_subset, recent_concussion == FALSE)

summary(rpq13_lm <-lm(data = dt_long_concuss, RPQ13_CORRECTED/52 ~ 
                                           # BM7 + 
                                           age + 
                                           # ageStroopQscore + 
                                           # concussions,
                                           # covidPositive + 
                                           # dementia# + 
                                           # hadConcussion,
                                           # joke + 
                                           # memory + 
                                           # alcohol * 
                                           # alcoholAmount + 
                                           smoking 
                                           # MCI_TOT 
                                           # neuro_TOT + 
                                           # MH_TOT
                                   ))

rpq13_lm_dt = data.frame(cbind((cbind(Beta = coef(rpq13_lm), confint(rpq13_lm))), p_value = summary(rpq13_lm)$coefficients[,4]))

write.csv(rpq13_lm_dt,"dt_out/rpq13_lm_dt_smoking.csv", row.names=TRUE)
```

### Alcohol

```{r}
dt_noS_subset = subset(dt_noS, select = c(mostRecentConcussion,
                                           RPQ13_CORRECTED,
                                           # BM7 + 
                                           age,
                                           # ageStroopQscore + 
                                           # concussions,
                                           # covidPositive + 
                                           # dementia
                                           # hadConcussion,
                                           # joke + 
                                           # memory + 
                                           alcohol,
                                           alcoholAmount
                                           # smoking
                                           # MCI_TOT
                                           ))
dt_noS_subset = na.omit(dt_noS_subset)
dt_noS_subset$recent_concussion = with(dt_noS_subset, ifelse(mostRecentConcussion %in% 
                                                 c("New_User_ConcussionC_1", 
                                                   "New_User_ConcussionC_2"), TRUE, FALSE))

dt_long_concuss = subset(dt_noS_subset, recent_concussion == FALSE)

lvl.alcohol <- c("None","New_User_Alcohol_Amount_LT3","New_User_Alcohol_Amount_47","New_User_Alcohol_Amount_714","New_User_Alcohol_Amount_MT14")
lbl.alcohol <- c(0,1,2,3,4)
dt_long_concuss$alcohol_recoded <- ordered(dt_long_concuss$alcoholAmount, levels=lvl.alcohol, labels=lbl.alcohol)
dt_long_concuss$alcohol_recoded = as.numeric(dt_long_concuss$alcohol_recoded)

# dt_long_concuss$excessive_alcohol = with(dt_long_concuss, ifelse(alcoholAmount == "New_User_Alcohol_Amount_MT14", 
#                                                    TRUE, FALSE))

summary(rpq13_lm <-lm(data = dt_long_concuss, RPQ13_CORRECTED ~ 
                                           # BM7 + 
                                           age + 
                                           # ageStroopQscore + 
                                           # concussions,
                                           # covidPositive + 
                                           # dementia# + 
                                           # hadConcussion,
                                           # joke + 
                                           # memory + 
                                           # alcohol *
                                           alcoholAmount #+
                                           # smoking 
                                           # MCI_TOT 
                                           # neuro_TOT + 
                                           # MH_TOT
                      # alcohol_recoded
                                   ))

rpq13_lm_dt = data.frame(cbind((cbind(Beta = coef(rpq13_lm), confint(rpq13_lm))), p_value = summary(rpq13_lm)$coefficients[,4]))

write.csv(rpq13_lm_dt,"dt_out/rpq13_lm_dt_alcohol.csv", row.names=TRUE)
```


### MCI

```{r}
dt_noS_subset = subset(dt_noS, select = c(mostRecentConcussion,
                                           RPQ13_CORRECTED,
                                           # BM7 + 
                                           age,
                                           # ageStroopQscore + 
                                           # concussions,
                                           # covidPositive + 
                                           # dementia
                                           # hadConcussion,
                                           # joke + 
                                           # memory + 
                                           # alcohol,
                                           # alcoholAmount
                                           # smoking
                                           MCI_TOT
                                           ))

dt_noS_subset$recent_concussion = with(dt_noS_subset, ifelse(mostRecentConcussion %in% 
                                                 c("New_User_ConcussionC_1", 
                                                   "New_User_ConcussionC_2"), TRUE, FALSE))

dt_long_concuss = subset(dt_noS_subset, recent_concussion == FALSE)

summary(rpq13_lm <-lm(data = dt_long_concuss, RPQ13_CORRECTED/52 ~ 
                                           # BM7 + 
                                           age + 
                                           # ageStroopQscore + 
                                           # concussions,
                                           # covidPositive + 
                                           # dementia# + 
                                           # hadConcussion,
                                           # joke + 
                                           # memory + 
                                           # alcohol * 
                                           # alcoholAmount #+
                                           # smoking 
                                            MCI_TOT 
                                           # neuro_TOT + 
                                           # MH_TOT
                                   ))

rpq13_lm_dt = data.frame(cbind((cbind(Beta = coef(rpq13_lm), confint(rpq13_lm))), p_value = summary(rpq13_lm)$coefficients[,4]))

write.csv(rpq13_lm_dt,"dt_out/rpq13_lm_dt_MCI.csv", row.names=TRUE)
```

### Plot

Note: do not show dementia because it's hard to explain
```{r}
plot.mci = read.csv("dt_out/rpq13_lm_dt_MCI.csv")
plot.alcohol = read.csv("dt_out/rpq13_lm_dt_alcohol.csv")
plot.smoking = read.csv("dt_out/rpq13_lm_dt_smoking.csv")
# plot.mci = read.csv("dt_out/rpq13_lm_dt_dementia.csv")
plot.sum = rbind(plot.smoking,plot.mci)
plot.sum = rbind(plot.sum,plot.alcohol)

plot.sum$significance = "p-value > 0.05"
plot.sum$significance[plot.sum$p_value < 0.05] <- "p-value < 0.05"

plot.sum = subset(plot.sum, X != "(Intercept)")
plot.sum = subset(plot.sum, X != "age")

ggplot2::ggplot(plot.sum, 
       aes(x=X, y=Beta/52, color = p_value, fill = p_value)) + 
    geom_point(shape=21, size = 2.5) +
    geom_errorbar(aes(ymin=X2.5../52, ymax=X97.5../52),
                  width=0,                    # Width of the error bars
                  size = 1.2,
                  position=position_dodge(.9)) +
  scale_color_gradient(
  low = "#ED2024",
  high = "#939598") + 
  scale_fill_gradient(
  low = "#ED2024",
  high = "#939598") + 
  theme_classic() + 
  # geom_hline(yintercept = 1, linetype="dotted") +
  coord_flip()+ylab("Beta") + 
  xlab("") 
ggsave("fig/long_concussion_rpq13_lm_MCI_smoking_alcohol.pdf", width = 7, height = 3,  bg = "transparent")
```



### Decision tree
```{r eval = FALSE}
tr1 <- trainControl(
        method = 'LGOCV',
        number = 1,
        p = 0.75,
        savePredictions = TRUE)
m1 <- train(RPQ13_CORRECTED ~ .,
        data = dt_noS_rpq_tot_na,
        method = 'rpart2',
        tuneLength = 10,
        metric = 'MAE',
        trControl = tr1)
plot(m1)
```

```{r eval = FALSE}
rpart.plot(m1$finalModel)
```








